\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-04-17 main.tex
%>% to |>

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat %>%
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long %>% face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% unfold(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% face_spatial() %>% face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat %>%
+   tsibble::as_tsibble(key = id, index = date)
R> 
R> dt <-  raw %>%
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt %>%
+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(sf)
R> # create a cubble
R> cb <- climate_flat %>%
+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
R> 
R> aus <- ozmaps::abs_ste
R> 
R> dt <- cb %>%
+   mutate(
+     # create `sfc` column based on long and lat
+     ll = st_sfc(
+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
+       crs = st_crs(aus)),
+ 
+     # append state multi-polygon based on the `sfc` created
+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+ 
+     # adopt a different projection: lambert conformal conic (EPSG:3112)
+     state = st_transform(state, crs = "EPSG:3112")
+     )
R> 
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name   wmo_id ts                        ll
  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
# ... with 1 more variable: state <MULTIPOLYGON [m]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(s2)
R> # Western Australia map
R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
R> 
R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
R> cb %>%
+   mutate(ll = s2_lnglat(long, lat)) %>%
+   filter(s2_within(ll, wa))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [3]: nested form
# bbox:     [115.97, -32.94, 117.18, -31.89]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts                 ll      
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- weatherdata::historical_tmax %>%
+   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% 
+   face_temporal() %>%
+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>%
+   group_by(month = lubridate::month(date), 
+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
+                                   "2016 ~ 2020", "1971 ~ 1975"))) %>%
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> nsw_vic <- ozmaps::abs_ste %>% 
+   filter(NAME %in% c("Victoria", "New South Wales"))
R> 
R> ggplot() + 
+   geom_sf(data = nsw_vic, 
+           fill = "transparent", color = "grey", linetype = "dotted") + 
+   geom_glyph(data = tmax, 
+              aes(x_major = longitude, x_minor = month, 
+                  y_major = latitude, y_minor = tmax,
+                  group = interaction(id, group), color = group),
+              width = 1, height = 0.5) +
+   ...
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> station_nested <- weatherdata::climate_full %>% 
+   mutate(cluster = ...)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cluster_nested <- station_nested %>% switch_key(cluster) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cluster_nested <- cluster_nested %>% get_centroid()
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # created SharedData instance for crosstalk
R> nested <- clean %>% SharedData$new(~id, group = "cubble")
R> long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
R> 
R> # create the spatial and temporal view each with a ShareData instance
R> p1 <- nested %>% ...
R> p2 <- long %>% ...
R> 
R> # Combine p1 and p2
R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # use the long form to create subplots for each station
R> df_id <- unique(clean$id)
R> p <- map(1:length(df_id), function(i){
+   dt <- clean %>% filter(id == df_id[i])
+   ggplot(dt) %>% ...
+ })
R> 
R> # create nested form leaflet map with temperature band as subplots 
R> nested <- face_spatial(clean)
R> leaflet(nested) %>% 
+   addTiles() %>% 
+   addCircleMarkers(group = "a", ...) %>% 
+   leafpop::addPopupGraphs(graph = p, ...)
\end{CodeInput}
\end{CodeChunk}

>>>>

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_nested <- cubble::climate_flat |>
+   as_cubble(key = id, index = date, coords = c(long, lat))
R> cubble_nested
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested |> face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long |> face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long |> unfold(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig |> face_spatial() |> face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # example with a tsibble created from climate_flat
R> raw <- climate_flat |>
+   tsibble::as_tsibble(key = id, index = date)
R> 
R> dt <-  raw |>
+   cubble::as_cubble(coords = c(long, lat))
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tbl_ts [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tbl_ts [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tbl_ts [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tbl_ts [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tbl_ts [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # add station-based features in the nested form.
R> dt |>
+   mutate(fabletools::features(ts, tmax, list(tmax_mean = mean, tmax_var = var)))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts       tmax_mean tmax_var
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>       <dbl>    <dbl>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tbl_ts>      25.7    38.6 
2 ASN00010311 -31.9  117. 179   york           94623 <tbl_ts>      26.2    51.1 
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tbl_ts>      23.7    45.4 
4 ASN00014015 -12.4  131.  30.4 darwin airpo~  94120 <tbl_ts>      33.1     3.02
5 ASN00015131 -17.6  134. 220   elliott        94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(sf)
R> # create a cubble
R> cb <- climate_flat |>
+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
R> 
R> aus <- ozmaps::abs_ste
R> 
R> dt <- cb |>
+   mutate(
+     # create `sfc` column based on long and lat
+     ll = st_sfc(
+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
+       crs = st_crs(aus)),
+ 
+     # append state multi-polygon based on the `sfc` created
+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+ 
+     # adopt a different projection: lambert conformal conic (EPSG:3112)
+     state = st_transform(state, crs = "EPSG:3112")
+     )
R> 
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name   wmo_id ts                        ll
  <chr>       <dbl> <dbl> <dbl> <chr>   <dbl> <list>           <POINT [°]>
1 ASN00009021 -31.9  116.  15.4 perth~  94610 <tibble> (115.9764 -31.9275)
2 ASN00010311 -31.9  117. 179   york    94623 <tibble>  (116.765 -31.8997)
3 ASN00010614 -32.9  117. 338   narro~  94627 <tibble> (117.1797 -32.9342)
4 ASN00014015 -12.4  131.  30.4 darwi~  94120 <tibble> (130.8925 -12.4239)
5 ASN00015131 -17.6  134. 220   ellio~  94236 <tibble> (133.5407 -17.5521)
# ... with 1 more variable: state <MULTIPOLYGON [m]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library(s2)
R> # Western Australia map
R> wa <- ozmaps::abs_ste |> filter(NAME == "Western Australia")
R> 
R> # mutate a `s2_lnglat` vector on `cb` created in the last chunk
R> cb |>
+   mutate(ll = s2_lnglat(long, lat)) |>
+   filter(s2_within(ll, wa))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [3]: nested form
# bbox:     [115.97, -32.94, 117.18, -31.89]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts                 ll      
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- weatherdata::historical_tmax |>
+   filter(between(stringr::str_sub(id, 7, 8), 46, 90))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |> 
+   face_temporal() |>
+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |>
+   group_by(month = lubridate::month(date), 
+          group = as.factor(ifelse(lubridate::year(date) > 2015, 
+                                   "2016 ~ 2020", "1971 ~ 1975"))) |>
+   summarise(tmax = mean(tmax, na.rm = TRUE))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |> face_spatial() |> filter(nrow(ts) == 24) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax |> face_temporal() |> unfold(latitude, longitude)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> nsw_vic <- ozmaps::abs_ste |> 
+   filter(NAME %in% c("Victoria", "New South Wales"))
R> 
R> ggplot() + 
+   geom_sf(data = nsw_vic, 
+           fill = "transparent", color = "grey", linetype = "dotted") + 
+   geom_glyph(data = tmax, 
+              aes(x_major = longitude, x_minor = month, 
+                  y_major = latitude, y_minor = tmax,
+                  group = interaction(id, group), color = group),
+              width = 1, height = 0.5) +
+   ...
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> station_nested <- weatherdata::climate_full |> 
+   mutate(cluster = ...)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cluster_nested <- station_nested |> switch_key(cluster) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cluster_nested <- cluster_nested |> get_centroid()
\end{CodeInput}
\end{CodeChunk}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\CommentTok{\# created SharedData instance for crosstalk}
\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\CommentTok{\# Combine p1 and p2}
\NormalTok{crosstalk}\SpecialCharTok{::}\FunctionTok{bscols}\NormalTok{(plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p1), plotly}\SpecialCharTok{::}\FunctionTok{ggplotly}\NormalTok{(p2), ...)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\CommentTok{\# use the long form to create subplots for each station}
\NormalTok{df\_id }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(clean}\SpecialCharTok{$}\NormalTok{id)}
\NormalTok{p }\OtherTok{\textless{}{-}} \FunctionTok{map}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(df\_id), }\ControlFlowTok{function}\NormalTok{(i)\{}
\NormalTok{  dt }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{==}\NormalTok{ df\_id[i])}
  \FunctionTok{ggplot}\NormalTok{(dt) }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
\NormalTok{\})}

\CommentTok{\# create nested form leaflet map with temperature band as subplots}
\NormalTok{nested }\OtherTok{\textless{}{-}} \FunctionTok{face\_spatial}\NormalTok{(clean)}
\FunctionTok{leaflet}\NormalTok{(nested) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{addTiles}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{addCircleMarkers}\NormalTok{(}\AttributeTok{group =} \StringTok{"a"}\NormalTok{, ...) }\SpecialCharTok{|\textgreater{}}
\NormalTok{  leafpop}\SpecialCharTok{::}\FunctionTok{addPopupGraphs}\NormalTok{(}\AttributeTok{graph =}\NormalTok{ p, ...)}
\end{Highlighting}
\end{Shaded}
